 Getting and Cleaning Data Course Project
========================================================
###Modified from http://mathijsdevaan.com/archives/227 (Mathijs de Vaan/md2999@columbia.edu)
# File-Name:       download_and_unzip.R           
# Date:            06_13_2014                                
# Author:          Jeff Roberts
# Email:           uwhawk@uw.edu                                      
# Purpose:         download and unzip files for the web
# Packages Used:   see load libraries and data 
# Machine:         Roberts' Dell laptop
# Github Path:     https://github.com/uwhawk/Getting_and_Cleaning_Data_Course_Project.git

Download and unzip the files from https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip 

```{r}
# load packages
require("R.utils")

# set workdrive
setwd("C:/R/zip")

# open a temporary directory to store data
temp <- tempdir()

# define the url  
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"

file <- basename(url)
  
# download the zip file and store in the "file" object 
download.file(url, file)

# unzip the file and store in the temp folder
gunzip(file, exdir = temp)
(l = list.files(temp))
# unlink the temporary directory
unlink(temp)
```

move working files from: C:\R\zip\getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip\UCI HAR Dataset\test & C:\R\zip\getdata###%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip###\UCI HAR Dataset\train from download to working directory: C:\R\zip using Windows Explorer.

```{r}
setwd("C:/R/zip")
test<- as.matrix(read.table("X_test.txt", header=FALSE))
train<-as.matrix(read.table("X_train.txt", header=FALSE))
###open row labels in dataframe
test.row<- as.matrix(read.table("y_test.txt", header=FALSE))
train.row<- as.matrix(read.table("y_train.txt", header=FALSE))
```

Assign Row Labels
```{r}
rownames(test) <- test.row[,1]
rownames(train) <- train.row[,1]
```


open features.txt which contains Column Lablels and descriptive names
```{r}
features<-as.data.frame(read.table("features.txt", header=FALSE))

features$V2<-gsub("mean()", "Mean Value",features$V2,ignore.case = TRUE)
features$V2<-gsub("std()", " Standard Deviation",features$V2,ignore.case = TRUE)
features$V2<-gsub("mad()", " Median absolute deviation",features$V2,ignore.case = TRUE)
features$V2<-gsub("max()", "Largest Value in array", features$V2,ignore.case = TRUE)
features$V2<-gsub("min()", "Smallest Value in array", features$V2,ignore.case = TRUE)
features$V2<-gsub("sma()", "Signal magnitude area", features$V2,ignore.case = TRUE)
features$V2<-gsub("energy()", "Energy measure. Sum of he squares divided by the number of values.", features$V2,ignore.case = TRUE)
features$V2<-gsub("iqr()", "Interquartile range", features$V2,ignore.case = TRUE)
features$V2<-gsub("entropy()", "Signal entropoy", features$V2,ignore.case = TRUE)
features$V2<-gsub("arCoeff()", "Autoregresion coefficients with Burg order equal to 4", features$V2,ignore.case = TRUE)
features$V2<-gsub("correlation()", "correlation coefficient between two signals",features$V2,ignore.case = TRUE)
features$V2<-gsub("maxInds()", "Index of the frequency component with largest magnitue",features$V2,ignore.case = TRUE)
features$V2<-gsub("meanFreq()", "Weighted average of the frequncy components to obtain a mean frequency",features$V2,ignore.case = TRUE)
features$V2<-gsub("skewness()", "Skewness of the frequency domain signal",features$V2,ignore.case = TRUE)
features$V2<-gsub("kurtosis()", "Kurtosis of the frequency domain signal",features$V2,ignore.case = TRUE)
features$V2<-gsub("bandsEnergy()", "Energy of frequency interval within the 64 bins of the FFT of each window",features$V2,ignore.case = TRUE)
features$V2<-gsub("angle()", "Angle between two vectors",features$V2,ignore.case = TRUE)
```

Assign Descriptive Column Labels to Training and Test Databases
```{r}
colnames(test) <- features[,2]
colnames(train) <- features[,2]
```

Combine test and training sets into a single dataset, add row labels as a column "user" to Subtotal Column Averages on.
```{r}
single<-as.data.frame(rbind(test, train))
a<-as.data.frame(as.numeric(rownames(single)))
colnames(a)<-"user"
single<-cbind(single,a)
```

Extract mean and standard deviation measurements for Tidy Dataset, compare dimentions of the datasets before and after mean and standard deviation extraction
```{r}
single.ms<-single[,grep("mean|Mean|standard|Standard|user", colnames(single), value=TRUE)]
dim(single)
dim(single.ms)
colnames(single)
colnames(single.ms)
```

Create second dataset with average variables (means/sd's) by user/subject
```{r}
(single.tidy<-aggregate(. ~ user, data = single.ms, mean))
dim(single.tidy)
```

write single.tidy to R report table and .csv files
```{r}
write.table(single.tidy, "tidy_data.dat", col.names = TRUE, row.names=TRUE)
write.csv(single.tidy, "tidy.csv", row.names=FALSE)
```